Sustainable Pathways for Irish homes in changing TIMES

### 2021-06-25 manuscript.tex
Done for 25-June-2021

In Ireland, previous modelling on the residential sector has often involved bottom-up residential models, rather than energy system models specifically examining the residential sector. Nevertheless, the residential models have provided insights on input data and methodology approach used in Ireland. 
Early work by Healy and CLinch ---
2009 Hull

@@

Mac Uidhir et al.\cite{Uidhir2020ImprovingDecisions}  created Archetype Dwelling Energy Model - Structured Query Language (ArDEM-SQL), to simulate the potential for improved energy efficiency gains within the existing retrofit programs, specifically the Better Energy Homes (BEH) scheme.
BER database had grown to in ex- cess of 735,0 0 0 records
We introduce a new methodology that provides insights into sub-optimal retrofit choices.
The five most common retrofit combinations are simulated for each of the nine archetypes. The 3 energy rating types and 3 archetypes
The results show that the alternative retrofit combination differs by archetype and that additional energy effi- ciency gains of up to 86% can be achieved due to alternative retrofit choices
while maintaining the total number of retrofits completed, we estimate an energy savings potential of 1389 GWh, realising an additional 643 GWh or an 86% additional energy savings potential relative to those in the baseline data, see Fig. 7
The simulation of alternative archetype retrofit measures
yielded additional energy savings of 86% when compared with the observed energy savings as they appear in the baseline sce- nario. The additional energy savings are not evenly distributed across all building archetypes with detached CD/EFG and terraced CD/EFG account for 98% of the potential additional energy sav- ings when using the alternative retrofits identified, suggesting that greater energy savings are possible when retrofit measures are ap- plied to less energy efficient dwellings. It
\citep{Ahern2020ADatabases}
and some models have been used to optimise Ireland's entire energy system \cite{Chiodi2013ModellingSystem}\cite{Connolly2021AalborgEnergy-System}, but very few models have used energy system optimisation model to explore the Irish residential sector. In 2014, Rogan et al used OSeMOYOS to explore energy service demands \cite{Rogan2014LEAPsSystem}, and Thellufsen examining DH with EnergyPLAN and TIMES in 2019 \cite{Thellufsen2019ImplementingIreland}.

@@

The ArDEM-SQL is a simulation model built on IS EN 13790, it converts the BER public database into Microsoft SQL language and then calculates annual energy consumption for space heating, water heating, pumps \& fans and lighting. ArDEM-SQL is used to clean the input data in TIM, by changing internal temperatures. The internal temperature is highly dependent on occupancy behaviour, a thermal comfort  thermal discomfort are higher among the over-65s than for other household \cite{Healy2002FuelIreland} - living room temperatues between 12 and 26...s  and therefore there is are large variations per BER rating.
The International Standard ISO 7730 [6] uses these PMV and PPD indices to predict the thermal sensation of people exposed to moderate thermal environments, as well as to specify acceptable thermal environmental conditions for comfort...
Despite this, it is
worth noting that the most inefficient dwellings i.e. band F, managed to achieve an
average temperature during period 2 that was only 0.5°C lower than the most
efficient dwellings i.e. band C.
C = 19.2, D = 18, E = 18.2 and F= 18.7 older occupants dublin Ho
%Healy & Clinch reported that fuel-poor households were more likely to be experiencing colder temperatures than other households. In total almost 30% of fuelpoor households had living room temperatures below 18°C and more than two thirds had living room temperatures below 20°C. Overall nearly half of all non-fuel-poor homes had living room temperatures below 20°C. In households with persons aged over 65 years, 16% had living room temperatures below 18°C and over 50% had temperatures below 20°C. On the other end of the scale over 5% of households occupied by persons over 65 years experienced living room temperatures of ≥24°C.https://arrow.tudublin.ie/cgi/viewcontent.cgi?article=1089&context=scienmas
%The BER for the sample dwellings ranged from C1 to F. Over two thirds of the
sample dwellings had a BER of E or F. The remaining dwellings had a BER of C or
D. In total 4 of the sample dwellings had undergone energy efficiency upgrades in
recent years, including cavity wall insulation, attic insulation where applicable and
heating control upgrades. In the case of two sample dwellings, these upgrade
works were completed between monitoring period 1 and 2, and therefore results
could be compared before and after retrofit.

>>>>

%and some models have been used to optimise Ireland's entire energy system \cite{Chiodi2013ModellingSystem}\cite{Connolly2021AalborgEnergy-System}, but very

In 2014, Rogan et al. \cite{Rogan2014LEAPsSystem} used OSeMOYOS to explore residential energy service demands in Ireland
 LEAP writes an input file for OSeMOSYS, runs it and imports its results back into LEAP. Currently, the LEAP input file only contains the aggregated electricity demand as the main driver for the optimization within OSeMOSYS.
 the model examines the
 3 main sceanrios - exploring transtport, residential, services and industy. The reference sceanrio - 
REF - TFC increase by 6.7\% in 2020, base year is 2008. space and water heating TFC have decreased in share of total residential TFC from 82 to 77 \%, whereas lighting, cooking and appliances have increased in TFC from 18 to 23 \%. assume that 1.4 \% of existing dwellings per annum will undergo a shallow retrofit in the reference
case, corresponding to just below 300,000 dwellings

In the energy efficiency scenario, we assume that 800,000 residential dwellings will be retrofitted between 2009 and 2020, in line with government targets for the BEH scheme,  Other policies in the residential sector are building regulations 2011 and building regulations 2013 and CFL lighting,  The combined impact of these policies is a reduction in TFC of 9.9 \% compared to the reference scenario.

The energy efficiency+scenario explores the impact of an ambitious retrofitting scenario beyond its current rate of implementation: retrofitbest is realised by retrofitting the same overall number of dwellings but undertaking a deeper, more comprehensive package of
energy efficiency measures on each, resulting in 7.2 \% savings compared to the energy efficiency scenario and 16.5 \% savings compared to the reference scenario.
  Insights - A challenging issue for energy modelling of policy is the
barriers and obstacles to investment, particularly in
residential energy efficiency measures such as split
incentives and information gaps, which lead to suboptimal
uptake of economically beneficial energy
efficiency improvements.

Thellufsen examining DH with EnergyPLAN and TIMES in 2019 \cite{Thellufsen2019ImplementingIreland}.

In Ireland few models have used energy system optimisation model to explore the residential sector. The majority of previous modelling on the residential sector in Ireland has involved bottom-up residential sector only models, rather than energy system models specifically examining the residential sector. Nevertheless, the residential models have provided insights on input data and methodology approach used in Ireland, some of which are outlined below.

Early work by Healy and CLinch ---\cite{Clinch2000HousingMortality,Healy2002FuelIreland,Clinch2003ValuingModel,Clinch2000DomesticFailure,Clinch2001Cost-benefitEfficiency}

@@

The Archetype Dwelling Energy Model (ArDEM-SQL) was developed by Mac Uidhir et al.\cite{Uidhir2020ImprovingDecisions}. ArDEM-SQL converts the Irish BER public database into Microsoft SQL language and then calculates annual energy consumption to simulate the potential for improved energy efficiency gains within the existing retrofit programs. Mac Uidhir et al. specifically examined the Better Energy Homes (BEH) scheme. The five most common retrofit combinations are simulated for each of 9 archetypes ( 3 energy rating types and 3 building types ). The results show that the alternative retrofit scenarios can achieve additional energy efficiency gains of up to 86\% can be achieved when compared the baseline scenario. This shows the the type of retrofit and type of house can have a huge affect on the energy saved. 
Greater energy savings are possible when retrofit measures are applied to less energy efficient dwellings (E, F and G-ratings ), albeit the ``rebound effect" was not accounted for due to insufficient data on the scale of the rebound effect and the relative difference between the two scenarios examined is considered negligible.

SEAI’s National Energy Modelling Framework (NEMF) is a suite of modelling tools used for modelling the Irish
energy system, both supply and demand side, to inform Irish Government energy policy and project Ireland’s
energy use based on these policie - https://www.seai.ie/publications/National-Energy-Projections-to-2030.pdf
The macro-economic model currently in use is called COSMO (Core Structural Model of the Irish Economy),
which creates a baseline forecast of energy consumption by sector (ESRI, 2017c). A range of sectoral models
are used to analyse both supply and demand side energy. Electricity supply is modelled using a commercial
power system software called PLEXOS47. A BioHeat model is used for techno-economic modelling of
bioenergy supply chains in Ireland for the electricity generation, transport and heat sectors48. There is also a
separate transport energy simulation model, which considers biofuels blending and the growth of low
emissions vehicles in the vehicle stock, including electric vehicles. In addition, separate modelling by an
energy-efficiency policy analysis tool (Unlocking model) informed the energy efficiency estimates included in
the NEMF49.

@@

The ArDEM-SQL, which was described in Section \ref{Intro:ESOMIreland} is a simulation model built on IS EN 13790. ArDEM-SQL is used to clean the input data in TIM, by changing internal temperatures. The internal temperature is highly dependent on occupancy behaviour, a thermal comfort  thermal discomfort are higher among the over-65s than for other household \cite{Healy2002FuelIreland} - living room temperatues between 12 and 26...s  and therefore there is are large variations per BER rating.
The International Standard ISO 7730 [6] uses these PMV and PPD indices to predict the thermal sensation of people exposed to moderate thermal environments, as well as to specify acceptable thermal environmental conditions for comfort...
Despite this, it is
worth noting that the most inefficient dwellings i.e. band F, managed to achieve an
average temperature during period 2 that was only 0.5°C lower than the most
efficient dwellings i.e. band C.
C = 19.2, D = 18, E = 18.2 and F= 18.7 older occupants dublin Ho
%Healy & Clinch reported that fuel-poor households were more likely to be experiencing colder temperatures than other households. In total almost 30% of fuelpoor households had living room temperatures below 18°C and more than two thirds had living room temperatures below 20°C. Overall nearly half of all non-fuel-poor homes had living room temperatures below 20°C. In households with persons aged over 65 years, 16% had living room temperatures below 18°C and over 50% had temperatures below 20°C. On the other end of the scale over 5% of households occupied by persons over 65 years experienced living room temperatures of ≥24°C.https://arrow.tudublin.ie/cgi/viewcontent.cgi?article=1089&context=scienmas
%The BER for the sample dwellings ranged from C1 to F. Over two thirds of the
sample dwellings had a BER of E or F. The remaining dwellings had a BER of C or
D. In total 4 of the sample dwellings had undergone energy efficiency upgrades in
recent years, including cavity wall insulation, attic insulation where applicable and
heating control upgrades. In the case of two sample dwellings, these upgrade
works were completed between monitoring period 1 and 2, and therefore results
could be compared before and after retrofit.
